---
title: "Clustering of DAT cells"
author: "Davide Risso"
date: '`r Sys.Date()`'
output: 
  html_document:
    fig_height: 7
    fig_width: 7
    toc: yes
    code_folding: hide
    toc_float: yes
---

```{r options, include=FALSE}
knitr::opts_chunk$set(cache=TRUE, error=FALSE, message=FALSE,
                      warning=FALSE, resuts = "hide")

library(scone)
library(stringi)
library(RColorBrewer)
library(clusterExperiment)
library(Rtsne)
library(matrixStats)

cols <- c(brewer.pal(9, "Set1"))

options(getClass.msg=FALSE)
NMF::nmf.options(grid.patch=TRUE)
```

```{r datain}
load("data/Oct15/imageAfterFiltering.RData")
dat <- sf.sc.eSet[,sf.sc.eSet$MD_expt_condition == "DAT_EXPT1"]
qc <- protocolData(dat)@data[,-c(6:9, 19)]

batch <- droplevels(dat$MD_c1_run_id)
chip_size <- dat$MD_c1_chip
```

# Filtering low-quality cells

```{r filtering}
counts <- na.omit(assayData(dat)$counts_table)
counts <- counts[rowSums(counts)>0,]

num_reads <- quantile(counts[counts > 0])[4]
num_cells <- 0.25*ncol(counts)
is_common <- rowSums(counts >= num_reads ) >= num_cells
table(is_common)

data(housekeeping)
hk <- intersect(stri_trans_totitle(housekeeping$V1), rownames(counts))

mfilt <- metric_sample_filter(counts,
                             nreads = qc$NREADS,
                             ralign = qc$RALIGN,
                             hard_nreads = 500000,
                             hard_ralign = 85,
                             gene_filter = is_common,
                             pos_controls = rownames(counts) %in% hk,
                             zcut = 3, mixture = FALSE,
                             plot = TRUE)

# Mean log10(x+1) expression
mu_obs <- rowMeans(log10(counts[hk,]+1))

# Assumed False Negatives
drop_outs <- counts[hk,] == 0

# Logistic Regression Model of Failure
ref.glms = list()
for (si in 1:dim(drop_outs)[2]){
  fit = glm(cbind(drop_outs[,si],1 - drop_outs[,si]) ~ mu_obs,
            family=binomial(logit))
  ref.glms[[si]] = fit$coefficients
}

# Plot Failure Curves and Calculate AUC
plot(NULL, main = "False Negative Rate Curves",
     ylim = c(0,1),xlim = c(0,6), 
     ylab = "Failure Probability", xlab = "Mean log10 Expression")
x = (0:60)/10
AUC = NULL
for(si in 1:ncol(counts)){
  y = 1/(exp(-ref.glms[[si]][1] - ref.glms[[si]][2] * x) + 1)
  AUC[si] = sum(y)/10
  lines(x, 1/(exp(-ref.glms[[si]][1] - ref.glms[[si]][2] * x) + 1),
        type = 'l', lwd = 2, col = cols[as.numeric(batch)[si]+1])
}

filtersample = !apply(simplify2array(mfilt[!is.na(mfilt)]), 1, any)
table(filtersample)

#filtered <- counts[,filtersample]
filtergenes <- rowSums(counts>10)>=10
table(filtergenes)
filtered <- counts[filtergenes,]
qc <- qc[colnames(filtered),]

pca <- prcomp(t(log1p(filtered)), scale. = TRUE)
plot(pca$x[,1:2], pch=19, col=cols[as.numeric(filtersample)+1], main="Raw data")
legend("bottomright", c("low-quality", "high-quality"), fill=cols)

fq <- FQ_FN(filtered)
pca <- prcomp(t(log1p(fq)), scale. = TRUE)
plot(pca$x[,1:2], pch=19, col=cols[as.numeric(filtersample)+1], main="FQ data")
legend("bottomright", c("low-quality", "high-quality"), fill=cols)

qcpca <- prcomp(qc, scale. = TRUE)
plot(qcpca$x[,1:2], pch=19, col=cols[as.numeric(filtersample)+1], main="QC PCA")
legend("bottomright", c("low-quality", "high-quality"), fill=cols)

plot(qc[,c(1, 3)], pch=19, col=cols[as.numeric(filtersample)+1])
legend("bottomright", c("low-quality", "high-quality"), fill=cols)
```

For now, I am removing all the low-quality cells. But I'm wondering if we should
revisit this. Could they be biologically different cell types that express fewer
genes? This leaves us with `r sum(filtersample)` cells out of `r length(filtersample)`.

```{r filtered}
filtered <- counts[,filtersample]
filtergenes <- rowSums(filtered>10)>=10
filtered <- filtered[filtergenes,]
qc <- qc[colnames(filtered),]
batch <- droplevels(batch[filtersample])
chip_size <- droplevels(chip_size[filtersample])

pca <- prcomp(t(log1p(filtered)), scale. = TRUE)
plot(pca$x[,1:2], pch=19, col=cols[batch], main="Raw data")

fq <- FQ_FN(filtered)
pca <- prcomp(t(log1p(fq)), scale. = TRUE)
plot(pca$x[,1:2], pch=19, col=cols[batch], main="FQ data")

qcpca <- prcomp(qc, scale. = TRUE)
plot(qcpca$x[,1:2], pch=19, col=cols[batch], main="QC PCA")

plot(qc[,c(1, 3)], pch=19, col=cols[batch], main="QC")
```

# Normalization

Here, we use `scone` to normalize the samples. We use housekeeping genes as negative control genes and a collection of positive controls from Poulin and Linnarson.

```{r controls}
## Positive controls: genes found DE between 6 DA neuron subtypes by Poulin et al
poulin <- c("Sox6", "Sncg", "Ndnf", "Igf1", "Foxa2", "Lmx1a", "Aldh1a1", "Slc32a1", "Satb1", "Clstn2", "Adcyap1", "Lpl", "Otx2", "Vip", "Chrna4", "Gsg1l", "Snca", "Ntf3")

linnarson <- c("Epha4", "Chrna5", "Nrip3", "Kcns3", "Cplx1", "Sox6", "Ndnf", "Kifc3", "Calb1", "Chst8", "Aldh1a1", "Igfbp2", "Lama5", "Anxa1", "Rpb4", "Aldh1a7", "Adcyap1", "Lhfpl2", "Cbln4", "Lpl", "Nhlh2", "Otx1", "Syn2", "Cbln1", "Gpx3", "Fjx1", "Foxa2", "En2", "Ntf3", "Gfra2", "Lix1", "Ptpn5", "Fgf1", "Nostrin", "Serpine2", "Kcnip3", "Grik1", "Lypd1", "Pou3f1", "Cd9", "Otx2", "Neurod6", "Grp", "Tcf12", "Calca", "Gpr83", "Vip", "Cck", "Cnr1", "Nphp1", "Chtf8", "Slc32a1", "Ctxn3", "Etv1", "Lmx1a")
poscon <- intersect(c(poulin, linnarson), rownames(fq))

## Negative controls
data(housekeeping)
hk <- intersect(rownames(fq), stri_trans_totitle(housekeeping[,1]))

plotHeatmap(log1p(fq[poscon,]), main="Positive controls")
plotHeatmap(log1p(fq[hk,]), main="Negative controls")
```

```{r scone}
set.seed(474)
negcon_ruv <- sample(hk, round(length(hk)/2))
negcon_eval <- setdiff(hk, negcon_ruv)

SUM_FN = function (ei)
{
  sums = colSums(ei)
  eo = t(t(ei)*mean(sums)/sums)
  return(eo)
}

scale_funs <- list(none=identity, # Identity - do nothing
             sum = SUM_FN,  # User-defined functions
             tmm = TMM_FN, # SCONE library wrappers...
             fq = FQT_FN,
             fq2 = FQ_FN)

scone_obj <- SconeExperiment(filtered, qc=qc, batch=batch,
                             negcon_ruv = rownames(filtered) %in% negcon_ruv,
                             negcon_eval = rownames(filtered) %in% negcon_eval,
                             poscon = rownames(filtered) %in% poscon)
scone_res <- scone(scone_obj, zero="postadjust", scaling=scale_funs,
                   adjust_batch="yes", adjust_bio="no",
                   bpparam=BiocParallel::MulticoreParam(7))
```


```{r scone_plots}
pc_obj <- prcomp(apply(t(get_scores(scone_res)),1,rank),
                center = TRUE,scale = FALSE)
bp_obj <- biplot_color(pc_obj,y = -get_score_ranks(scone_res),expand = .6)
points(t(bp_obj[1,]), pch = 1, col = "red", cex = 1)
points(t(bp_obj[1,]), pch = 1, col = "red", cex = 1.5)
points(t(bp_obj[rownames(bp_obj) == "none,none,no_uv,no_bio,no_batch",]),
       pch = 1, col = "blue", cex = 1)

if(FALSE) {
  biplot_interactive(scone_res)
}

head(get_scores(scone_res))

norm1 <- get_normalized(scone_res, "none,fq2,ruv_k=3,no_bio,batch")
norm2 <- get_normalized(scone_res, "none,fq,ruv_k=3,no_bio,no_batch")
norm3 <- get_normalized(scone_res, "none,fq2,no_uv,no_bio,no_batch")

plotHeatmap(log1p(norm1[poscon,]), main="none,fq2,ruv_k=3,no_bio,batch")
plotHeatmap(log1p(norm2[poscon,]), main="none,fq,ruv_k=3,no_bio,no_batch")
plotHeatmap(log1p(norm3[poscon,]), main="none,fq2,no_uv,no_bio,no_batch")

```

# Clustering

```{r tightcluster}
seed <- 927501

rsec <- RSEC(norm3, isCount = TRUE, minSizes = 5, alphas = 0.3,
             combineProportion = 0.5,
             mergeMethod = "adjP", mergeCutoff = 0.05,
             ncores = 7, random.seed = seed, run = TRUE)

# cl <- clusterMany(norm3, isCount = TRUE, dimReduce="PCA", nPCADims = 50, alphas = 0.3,
#                   sequential = TRUE, subsample = TRUE, minSizes = 5,
#                   clusterFun = c("hierarchical01", "tight"),
#                   ks = 4:15, random.seed = seed,
#                   ncores = 7)
# cl <- combineMany(cl, proportion = .5)
# cl <- makeDendrogram(cl, dimReduce = "mad", ndims = 1000)
# plotDendrogram(cl)
# cl <- mergeClusters(cl, mergeMethod = "adjP", cutoff = 0.05)
# plotClusters(cl)

plotClusters(rsec)
```

```{r final}
final_tight <- factor(clusterMatrixNamed(rsec)[,2], levels=clusterLegend(rsec)[[2]][, "name"])
names(final_tight) <- colnames(fq)

final_merged <- factor(primaryClusterNamed(rsec), levels=clusterLegend(rsec)[[1]][, "name"])
names(final_merged) <- colnames(fq)

print(table(final_tight, final_merged))

plotDendrogram(rsec)
plotCoClustering(rsec, whichClusters = 1:2)
```

# Visualization

```{r visualize}
wh_rm <- which(final_tight=="-1")
pca <- prcomp(t(log1p(norm3)[,-wh_rm]), scale. = TRUE, center = TRUE)

cols1 <- clusterLegend(rsec)[[2]][, "color"]
names(cols1) <- clusterLegend(rsec)[[2]][, "name"]
cols2 <- clusterLegend(rsec)[[1]][, "color"]
names(cols2) <- clusterLegend(rsec)[[1]][, "name"]

colFinal <- cols1[final_tight]
colMerged <- cols2[final_merged]

plot(pca$x, pch=19, col=colFinal[-wh_rm], main="PCA")
legend("topright", levels(final_tight), fill=cols1[levels(final_tight)])
plot(pca$x, pch=19, col=colMerged[-wh_rm], main="PCA, merged")
legend("topright", levels(final_merged), fill=cols2[levels(final_merged)])

vars <- rowVars(log1p(norm1))
names(vars) <- rownames(norm1)
vars <- sort(vars, decreasing = TRUE)

tsne_data <- Rtsne(t(log1p(norm3)[names(vars)[1:500],-wh_rm]),
                   max_iter=5000, perplexity = 10)
plot(tsne_data$Y, pch=19, col=colFinal[-wh_rm], main="t-SNE, 500 most variable genes")
plot(tsne_data$Y, pch=19, col=colMerged[-wh_rm], main="t-SNE, 500 most variable genes, merged")

plotHeatmap(rsec, whichClusters=1:2, clusterFeaturesData=poscon)
```

# Marker genes

```{r heatmap}
primaryClusterIndex(rsec) <- 2
genes <- getBestFeatures(rsec, contrastType = "Pairs", isCount=TRUE)
head(genes)

plotHeatmap(rsec, clusterSamplesData = "primaryCluster",
            clusterFeaturesData=unique(genes[,"IndexInOriginal"]))

plotHeatmap(rsec, clusterFeaturesData=unique(genes[,"IndexInOriginal"]))

```

```{r markers}
markers <- list(Set1=c("Erbb4", "Pcsk1", "Pepd", "Pdp2", "Nubp2", "Pik3c3", "Cacna2d3", "Kcns3", "Chrna3", "Epha4"),
                Set2=c("Adamts2", "Arpp21", "Rasgrp1", "Pitx2", "Irx6", "Cd36", "Pgr15l", "Gda", "Nos1", "Tac1"),
                Set3=c("Gad2", "Rasgrf1", "Satb2", "Fam159b", "Otof", "Msn", "Six6", "Gad1", "Ecel1", "Calcr"),
                Set4=c("Cnr1", "Cbln4", "Lpl", "Gpr83", "Neurod6", "Gkn1", "Igfbp4", "Prdm8", "Nxph3", "Tiam2")
)
markers <-lapply(markers, function(x) which(rownames(rsec) %in% x))
blank_data <- makeBlankData(transform(rsec), markers)

dat <- as.matrix(blank_data$dataWBlanks)
rnames <- blank_data$rowNamesWBlanks
rownames(dat) <- rnames
breaks <- unique(c(min(dat, na.rm = TRUE), seq(0, quantile(dat[dat > 0], .99, na.rm = TRUE), length = 50), max(dat, na.rm = TRUE)))

NMF::aheatmap(dat[,order(final_tight)], Rowv=NA, Colv=NA, color = seqPal5, annCol = data.frame(Clusters=sort(final_tight)), annColors = list(Clusters=cols1), breaks=breaks, main="Cluster markers")


linnarson <- list(SNC_VTA1=c("Epha4", "Chrna5", "Nrip3", "Kcns3", "Cplx1", "Sox6", "Ndnf"),
                  Not_SNC=c("Kifc3", "Calb1", "Chst8"),
                  SNC_VTA2=c("Aldh1a1", "Igfbp2", "Lama5", "Anxa1", "Rpb4", "Aldh1a7"),
                  VTA2_VTA3=c("Adcyap1", "Lhfpl2", "Cbln4", "Lpl", "Nhlh2", "Otx1"),
                  VTA3_VTA4=c("Syn2", "Cbln1", "Gpx3"),
                  VTA1_VTA3=c("Fjx1", "Foxa2", "En2", "Ntf3", "Gfra2"),
                  SNC=c("Lix1", "Ptpn5", "Fgf1", "Nostrin", "Serpine2", "Kcnip3", "Grik1"),
                  VTA1=c("Lypd1", "Pou3f1", "Cd9"),
                  VTA2=c("Otx2", "Neurod6", "Grp", "Tcf12", "Calca", "Gpr83"),
                  VTA3=c("Vip", "Cck", "Cnr1", "Nphp1", "Chtf8"),
                  VTA4=c("Slc32a1", "Ctxn3", "Etv1", "Lmx1a"))
markers <-lapply(linnarson, function(x) which(rownames(rsec) %in% x))
blank_data <- makeBlankData(transform(rsec), markers)

dat <- as.matrix(blank_data$dataWBlanks)
rnames <- blank_data$rowNamesWBlanks
rownames(dat) <- rnames
breaks <- unique(c(min(dat, na.rm = TRUE), seq(0, quantile(dat[dat > 0], .99, na.rm = TRUE), length = 50), max(dat, na.rm = TRUE)))

NMF::aheatmap(dat[,order(final_tight)], Rowv=NA, Colv=NA, color = seqPal5, annCol = data.frame(Clusters=sort(final_tight)), annColors = list(Clusters=cols1), breaks=breaks, main="Linnarson markers")
```

From the last heatmap, we can see that there is correspondence between our clusters
and Linnarson's clusters.

```{r info}
sessionInfo()
```
