---
title: "Clustering of DAT cells"
author: "Davide Risso"
date: '`r Sys.Date()`'
output: 
  html_document:
    fig_height: 7
    fig_width: 7
    toc: yes
    code_folding: hide
    toc_float: yes
---

```{r options, include=FALSE}
knitr::opts_chunk$set(cache=TRUE, error=FALSE, message=FALSE,
                      warning=FALSE, resuts = "hide")

library(scone)
library(stringi)
library(RColorBrewer)
library(clusterExperiment)

cols <- c(brewer.pal(9, "Set1"))

options(getClass.msg=FALSE)
NMF::nmf.options(grid.patch=TRUE)
```

```{r datain}
load("data/Oct15/imageAfterFiltering.RData")
dat <- sf.sc.eSet[,sf.sc.eSet$MD_expt_condition == "DAT_EXPT1"]
qc <- protocolData(dat)@data[,-c(6:9, 19)]

batch <- droplevels(dat$MD_c1_run_id)
chip_size <- dat$MD_c1_chip
```

# Filtering low-quality cells

```{r filtering}
counts <- na.omit(assayData(dat)$counts_table)
counts <- counts[rowSums(counts)>0,]

num_reads <- quantile(counts[counts > 0])[4]
num_cells <- 0.25*ncol(counts)
is_common <- rowSums(counts >= num_reads ) >= num_cells
table(is_common)

data(housekeeping)
hk <- intersect(stri_trans_totitle(housekeeping$V1), rownames(counts))

mfilt <- metric_sample_filter(counts,
                             nreads = qc$NREADS,
                             ralign = qc$RALIGN,
                             hard_nreads = 500000,
                             hard_ralign = 85,
                             gene_filter = is_common,
                             pos_controls = rownames(counts) %in% hk,
                             zcut = 3, mixture = FALSE,
                             plot = TRUE)

# Mean log10(x+1) expression
mu_obs <- rowMeans(log10(counts[hk,]+1))

# Assumed False Negatives
drop_outs <- counts[hk,] == 0

# Logistic Regression Model of Failure
ref.glms = list()
for (si in 1:dim(drop_outs)[2]){
  fit = glm(cbind(drop_outs[,si],1 - drop_outs[,si]) ~ mu_obs,
            family=binomial(logit))
  ref.glms[[si]] = fit$coefficients
}

# Plot Failure Curves and Calculate AUC
plot(NULL, main = "False Negative Rate Curves",
     ylim = c(0,1),xlim = c(0,6), 
     ylab = "Failure Probability", xlab = "Mean log10 Expression")
x = (0:60)/10
AUC = NULL
for(si in 1:ncol(counts)){
  y = 1/(exp(-ref.glms[[si]][1] - ref.glms[[si]][2] * x) + 1)
  AUC[si] = sum(y)/10
  lines(x, 1/(exp(-ref.glms[[si]][1] - ref.glms[[si]][2] * x) + 1),
        type = 'l', lwd = 2, col = cols[as.numeric(batch)[si]+1])
}

filtersample = !apply(simplify2array(mfilt[!is.na(mfilt)]), 1, any)

#filtered <- counts[,filtersample]
filtergenes <- rowSums(counts>10)>=10
table(filtergenes)
filtered <- counts[filtergenes,]
qc <- qc[colnames(filtered),]

pca <- prcomp(t(log1p(filtered)), scale. = TRUE)
plot(pca$x[,1:2], pch=19, col=cols[as.numeric(filtersample)+1], main="Raw data")
legend("bottomright", c("low-quality", "high-quality"), fill=cols)

fq <- FQ_FN(filtered)
pca <- prcomp(t(log1p(fq)), scale. = TRUE)
plot(pca$x[,1:2], pch=19, col=cols[as.numeric(filtersample)+1], main="FQ data")
legend("bottomright", c("low-quality", "high-quality"), fill=cols)

qcpca <- prcomp(qc, scale. = TRUE)
plot(qcpca$x[,1:2], pch=19, col=cols[as.numeric(filtersample)+1], main="QC PCA")
legend("bottomright", c("low-quality", "high-quality"), fill=cols)

plot(qc[,c(1, 3)], pch=19, col=cols[as.numeric(filtersample)+1])
legend("bottomright", c("low-quality", "high-quality"), fill=cols)
```

For now, I am removing all the low-quality cells. But I'm wondering if we should
revisit this. Could they be biologically different cell types that express fewer
genes?

```{r filtered}
filtered <- counts[,filtersample]
filtergenes <- rowSums(filtered>10)>=10
filtered <- filtered[filtergenes,]
qc <- qc[colnames(filtered),]
batch <- droplevels(batch[filtersample])
chip_size <- droplevels(chip_size[filtersample])

pca <- prcomp(t(log1p(filtered)), scale. = TRUE)
plot(pca$x[,1:2], pch=19, col=cols[batch], main="Raw data")

fq <- FQ_FN(filtered)
pca <- prcomp(t(log1p(fq)), scale. = TRUE)
plot(pca$x[,1:2], pch=19, col=cols[batch], main="FQ data")

qcpca <- prcomp(qc, scale. = TRUE)
plot(qcpca$x[,1:2], pch=19, col=cols[batch], main="QC PCA")

plot(qc[,c(1, 3)], pch=19, col=cols[batch])
```

# Normalization

Here, we normalize the samples by removing the batch effects and the first two QC PCs,
after FQ normalization


```{r norm}
Y <- t(log1p(fq))
W <- model.matrix(~batch + qcpca$x[,1:2])
alpha <- solve(t(W) %*% W) %*% t(W) %*% Y
correctedY <- t(Y - W %*% alpha)
#correctedY[correctedY<0] <- 0
pca_dat <- prcomp(t(correctedY), center=TRUE, scale=TRUE)

plot(pca_dat$x, pch=19, col=cols[batch])

## Positive controls: genes found DE between 6 DA neuron subtypes by Poulin et al
poulin <- c("Sox6", "Sncg", "Ndnf", "Igf1", "Foxa2", "Lmx1a", "Aldh1a1", "Slc32a1", "Satb1", "Clstn2", "Adcyap1", "Lpl", "Otx2", "Vip", "Chrna4", "Gsg1l", "Snca", "Ntf3")

linnarson <- c("Epha4", "Chrna5", "Nrip3", "Kcns3", "Cplx1", "Sox6", "Ndnf", "Kifc3", "Calb1", "Chst8", "Aldh1a1", "Igfbp2", "Lama5", "Anxa1", "Rpb4", "Aldh1a7", "Adcyap1", "Lhfpl2", "Cbln4", "Lpl", "Nhlh2", "Otx1", "Syn2", "Cbln1", "Gpx3", "Fjx1", "Foxa2", "En2", "Ntf3", "Gfra2", "Lix1", "Ptpn5", "Fgf1", "Nostrin", "Serpine2", "Kcnip3", "Grik1", "Lypd1", "Pou3f1", "Cd9", "Otx2", "Neurod6", "Grp", "Tcf12", "Calca", "Gpr83", "Vip", "Cck", "Cnr1", "Nphp1", "Chtf8", "Slc32a1", "Ctxn3", "Etv1", "Lmx1a")
poscon <- intersect(c(poulin, linnarson), rownames(fq))

plotHeatmap(log1p(fq[poscon,]))
plotHeatmap(correctedY[poscon,])
```

----> HERE <----

# Clustering

```{r tightcluster, eval=evalTight}
minSize<-10
mkFullClMat<-function(obj,rev=TRUE){ #little function for making clMat from the tightTracking output
  if(rev) obj$clMat<-rev(obj$clMat)
	clMat<-do.call("cbind",obj$clMat) 
	names(clMat)<-unlist(lapply(names(obj$clMat),function(x){paste(x,gsub("[.]","=",colnames(obj$clMat[[1]])),sep=",")}))
	return(clMat)
}

seed<-927501
beta<-0.9
ncores<-4

print(system.time(trackClusters_all<-trackCluster(pcaDat_dat,type = "tight", random.seed=seed,beta=beta,ncores=ncores)))
clMat<-mkFullClMat(trackClusters_all)
write.table(clMat,file="../data/dat_tightTracking_Oct15_filtered.txt",col.names=TRUE,sep="\t",row.names=FALSE)
```

```{r pullTight}
if(!evalTight){
  minSize<-10
  mkFullClMat<-function(obj,rev=TRUE){ #little function for making clMat from the tightTracking output
    if(rev) obj$clMat<-rev(obj$clMat)
    clMat<-do.call("cbind",obj$clMat) 
	  names(clMat)<-unlist(lapply(names(obj$clMat),function(x){paste(x,gsub("[.]","=",colnames(obj$clMat[[1]])),sep=",")}))
  	return(clMat)
  }

  seed<-927501
  beta<-0.9
  ncores<-4

  tmp<-read.table("../data/dat_tightTracking_Oct15_filtered.txt",header=TRUE,sep="\t")
  trackClusters_all<-list()
  trackClusters_all$clMat<-list("alpha=0.1"=tmp[,1:14],"alpha=0.2"=tmp[,15:28],"alpha=0.3"=tmp[,29:42])
}
```

```{r final, dependson="tightCluster", results="markup"}
options(expressions=1e5)
clMat<-mkFullClMat(trackClusters_all,rev=FALSE)[,-c(1, 3)] #remove the 'bad' parameter values. 
cl<-findSharedClusters(clMat,minSize=5,proportion=0.7)

out<-plotTracking(cbind(Pam=clusters_dat, ChooseFrom=clMat, chip_size, batch_dat))

print(table(clMat[,13], clusters_dat))
print(table(clMat[,28], clusters_dat))
print(table(clMat[,17], clusters_dat))

final_tight <- clMat[,17]
names(final_tight) <- colnames(fq_dat)

print(table(final_tight, chip_size))

write.table(final_tight,file="../data/dat_tightTracking_Final_Oct15_filtered.txt",row.names=colnames(fq_dat), sep="\t", col.names=FALSE)
```

```{r merge_clusters, dependson="final", results="markup"}
## use 1000 most variable genes to merge clusters
wh_rm <- which(final_tight == -1)

vars <- apply(correctedY_dat[,-wh_rm], 1, var)
vars <- sort(vars, decreasing=TRUE)

pca_v <- prcomp(t(correctedY_dat[names(vars)[1:1000],]), center = TRUE, scale. = TRUE)
merged <- clusterCells::mergeClusters(pca_v$x[,1:50], final_tight, mergeMethod = "adjP", plotType="mergeMethod")
final_merged <- as.factor(merged$cl)

table(final_merged, final_tight)
colFinal <- c("white", cc[1:10])
colFinal[colFinal==cc[9]] <- cc[12]
colMerged <- colFinal[c(1, 10, 8, 3, 9, 11, 5, 2, 7)]
colMergedNoWhite <- colMerged[-1]

barplot(table(final_tight, final_merged), col=colFinal, main="Merged clusters")
legend("topright", c("Unassigned", as.character(1:max(final_tight[-wh_rm]))), fill=colFinal)

pairs(pcaDat_dat[-wh_rm,1:3], pch=19, col=colMerged[final_merged[-wh_rm]], main="PCA of normalized counts")

pairs(pca_v$x[-wh_rm,1:3], pch=19, col=colMerged[final_merged[-wh_rm]], main="PCA of normalized counts (1,000 most variable genes)")

write.table(final_merged, file="DAT/cluster_labels_merged.txt", quote=FALSE, sep="\t", row.names=colnames(fq_dat))
```


```{r genesDE,dependson="final"}
whRm <- wh_rm <- which(final_tight=="-1")

## dendro <- clusterHclust(pcaDat_dat[-whRm,],final_tight[-whRm],full=FALSE)

dendro <- clusterHclust(pca_v$x[-wh_rm,1:npcs], final_tight[-wh_rm], full=FALSE)
# hGenesTable <- clusterCells::getBestGenes(cl=final_merged[-wh_rm], dat=t(correctedY_dat[,-wh_rm]), dendro=dendro, number=50, type="Dendro", p.value=0.05, returnType ="Table")
# 
# pairMat <- t(apply(expand.grid(levels(factor(final_merged[-wh_rm])), levels(factor(final_merged[-wh_rm]))), 1, sort))
# pairMat <- unique(pairMat)
# pairMat <- pairMat[-which(pairMat[,1]==pairMat[,2]),]
# hGenesPairsTable <- clusterCells::getBestGenes(cl=final_merged[-wh_rm], dat=t(correctedY_dat[,-wh_rm]), pairMat=pairMat, number=50, type="Pairs", p.value=0.05, returnType ="Table")
# 
# allGenes <- clusterCells::getBestGenes(cl=final_merged[-wh_rm], dat=t(correctedY_dat[,-wh_rm]), type="OneAgainstAll", contrastAdj="All", p.value=0.05, number=50)

hGenesTable <- clusterCells::getBestGenes(cl=final_tight[-wh_rm], dat=t(correctedY_dat[,-wh_rm]), dendro=dendro, number=50, type="Dendro", p.value=0.05, returnType ="Table")

pairMat <- t(apply(expand.grid(levels(factor(final_tight[-wh_rm])), levels(factor(final_tight[-wh_rm]))), 1, sort))
pairMat <- unique(pairMat)
pairMat <- pairMat[-which(pairMat[,1]==pairMat[,2]),]
hGenesPairsTable <- clusterCells::getBestGenes(cl=final_tight[-wh_rm], dat=t(correctedY_dat[,-wh_rm]), pairMat=pairMat, number=50, type="Pairs", p.value=0.05, returnType ="Table")
```

```{r heatmap,dependson="genesDE"}
dendro <- clusterHclust(pca_v$x[,1:npcs], final_tight, full=TRUE, unassigned="outgroup")
#dendro<-clusterHclust(pcaDat_dat,final_tight,full=TRUE,unassigned="outgroup")
par(mfrow=c(1,2))
outDendoHeat<-dualHeatmap(clusterVector=final_tight,heatData=t(fq_dat),clusterData=dendro,clusterSamples=TRUE,whVars=hGenesTable$Index,annCol=data.frame(FinalCluster=as.factor(final_tight),MergedCluster=final_merged, Chip_size=chip_size),main="Genes from Dendrogram Hierarchy",breaks=0.99, annColors=list(FinalCluster=colFinal,MergedCluster=colMerged, Chip_size=cc))
outPairsHeat<-dualHeatmap(clusterVector=final_tight,heatData=t(fq_dat),clusterData=dendro,clusterSamples=TRUE,whVars=unique(hGenesPairsTable$Index),annCol=data.frame(FinalCluster=as.factor(final_tight),MergedCluster=final_merged, Chip_size=chip_size),main="Genes from All Pairwise",breaks=0.99, annColors=list(FinalCluster=colFinal,MergedCluster=colMerged, Chip_size=cc))

pdf("DAT/heatmap_dendro.pdf",width=10,height=25)
outDendoHeat<-dualHeatmap(clusterVector=final_tight,heatData=t(fq_dat),clusterData=dendro,clusterSamples=TRUE,whVars=hGenesTable$Index,annCol=data.frame(FinalCluster=as.factor(final_tight),MergedCluster=final_merged, Chip_size=chip_size),main="Genes from Dendrogram Hierarchy",breaks=0.99, annColors=list(FinalCluster=colFinal,MergedCluster=colMerged, Chip_size=cc))
dev.off()
pdf("DAT/heatmap_pairs.pdf",width=10,height=25)
outPairsHeat<-dualHeatmap(clusterVector=final_tight,heatData=t(fq_dat),clusterData=dendro,clusterSamples=TRUE,whVars=unique(hGenesPairsTable$Index),annCol=data.frame(FinalCluster=as.factor(final_tight),MergedCluster=final_merged, Chip_size=chip_size),main="Genes from All Pairwise",breaks=0.99, annColors=list(FinalCluster=colFinal,MergedCluster=colMerged, Chip_size=cc))
dev.off()
```

```{r heatmap_curated, results="markup", dependson="genesDE"}
pdf("DAT/heatmap_curated.pdf")
dualHeatmap(clusterVector=final_tight,heatData=t(fq_dat),clusterData=dendro,clusterSamples=TRUE,whVars=gg,annCol=data.frame(FinalCluster=as.factor(final_tight),PAM=clusters_dat, Batch=batch_dat, Chip_size=chip_size),main="Manually Curated Genes", annColors=list(FinalCluster=colFinal, Chip_size=cc))
dev.off()
```

```{r writeGenes,dependson=c("genesDE","heatmap")}
write.table(hGenesTable[outDendoHeat$heatOut$rowInd,],file="DAT/DEGenes_byDendro_1024.txt",col.names=TRUE,row.names=FALSE,sep="\t",quote=FALSE)
write.table(hGenesPairsTable[outPairsHeat$heatOut$rowInd,],file="DAT/DEGenes_byPairs_1024.txt",col.names=TRUE,row.names=FALSE,sep="\t",quote=FALSE)
```

## Visualization

```{r tsne_grey, dependson="final"}
tsne_data <- read.table("DAT/tsne_all.txt")
idx <- apply(tsne_data, 1, function(x) max(abs(x))<=200)

colgrey <- rep(rgb(0.86, 0.86, 0.86, 0.5), nrow(tsne_data))
colgrey[expt=="DAT_EXPT1"] <- c(rgb(0, 0, 0, 0), colFinal[-1])[as.factor(final_tight)]

colgrey2 <- rep(rgb(0.86, 0.86, 0.86, 0.5), nrow(tsne_data))
colgrey2[expt=="DAT_EXPT1"] <- c(rgb(0, 0, 0, 0), colMerged[-1])[as.factor(final_merged)]

plot(tsne_data[idx,], t='p', main = "tSNE visualization (1,000 most variable genes)", col=colgrey[idx], pch = 19, xlab="Component 1", ylab="Component 2")

pdf("DAT/tsne_all_clusters.pdf")
plot(tsne_data[idx,], t='p', main = "tSNE visualization (1,000 most variable genes)", col=colgrey[idx], pch = 19, xlab="Component 1", ylab="Component 2")
dev.off()

pdf("DAT/tsne_all_clusters_merged.pdf")
plot(tsne_data[idx,], t='p', main = "tSNE visualization (1,000 most variable genes)", col=colgrey2[idx], pch = 19, xlab="Component 1", ylab="Component 2")
dev.off()

tsne_data <- read.table("DAT/tsne_all_500.txt")
idx <- apply(tsne_data, 1, function(x) max(abs(x))<=200)

plot(tsne_data[idx,], t='p', main = "tSNE visualization (500 most variable genes)", col=colgrey[idx], pch = 19, xlab="Component 1", ylab="Component 2")

pdf("DAT/tsne_all_500_clusters.pdf")
plot(tsne_data[idx,], t='p', main = "tSNE visualization (500 most variable genes)", col=colgrey[idx], pch = 19, xlab="Component 1", ylab="Component 2")
dev.off()

pdf("DAT/tsne_all_500_clusters_merged.pdf")
plot(tsne_data[idx,], t='p', main = "tSNE visualization (500 most variable genes)", col=colgrey2[idx], pch = 19, xlab="Component 1", ylab="Component 2")
dev.off()

tsne_data <- read.table("DAT/tsne_all_250.txt")
idx <- apply(tsne_data, 1, function(x) max(abs(x))<=200)

plot(tsne_data[idx,], t='p', main = "tSNE visualization (250 most variable genes)", col=colgrey[idx], pch = 19, xlab="Component 1", ylab="Component 2")

pdf("DAT/tsne_all_250_clusters.pdf")
plot(tsne_data[idx,], t='p', main = "tSNE visualization (250 most variable genes)", col=colgrey[idx], pch = 19, xlab="Component 1", ylab="Component 2")
dev.off()

pdf("DAT/tsne_all_250_clusters_merged.pdf")
plot(tsne_data[idx,], t='p', main = "tSNE visualization (250 most variable genes)", col=colgrey2[idx], pch = 19, xlab="Component 1", ylab="Component 2")
dev.off()
```

```{r pca_dat, dependson="final", results="markup"}
whRm <- which(final_tight == -1)
pdf("DAT/pca_tight_clustering.pdf")
pairs(pca_dat$x[-whRm,1:3], pch=20, col=cc[as.factor(final_tight[-whRm])], main="PCA")
## legend("bottomright", paste("Cluster", 1:max(final_tight)), fill=cc, cex=.5)
dev.off()

pdf("DAT/pca_qc_tight_clustering.pdf")
plot(qcpca_dat$x[-whRm,1:2], pch=20, col=cc[as.factor(final_tight[-whRm])], main="QC")
legend("bottomright", paste("Cluster", 1:max(final_tight)), fill=cc, cex=.5)
dev.off()

pdf("DAT/pca_nreads_tight_clustering.pdf")
plot(qc_dat[-whRm,c(1, 3)], pch=20, col=cc[as.factor(final_tight[-whRm])], main="QC")
legend("bottomright", paste("Cluster", 1:max(final_tight)), fill=cc, cex=.5)
dev.off()

## print(cumsum(pca_dat$sdev^2)/sum(pca_dat$sdev^2))

library(scatterplot3d)
pdf("DAT/pca_3d.pdf")
scatterplot3d(pca_dat$x[-whRm,1:3], main="3D Scatterplot", color=cc[as.factor(final_tight[-whRm])], pch=20)
dev.off()

```

```{r tsne, dependson="final"}
whRm <- which(final_tight == -1)
vars <- apply(correctedY_dat[,-whRm], 1, var)
vars <- sort(vars, decreasing=TRUE)

d <- dist(t(correctedY_dat[,-whRm][names(vars)[1:1000],]))

set.seed(123)
tsne_data <- tsne(d, k = 2, perplexity = 5, max_iter = 2000)
```

```{r tsne_plot, dependson=c("tsne", "final"), results="markup"}
idx <- apply(tsne_data, 1, function(x) max(abs(x))<=200)

plot(tsne_data[idx,], t='p', main = "tSNE visualization (1,000 most variable genes)", col=cc[as.factor(final_tight[-whRm][idx])], pch = 19, xlab="Component 1", ylab="Component 2")
legend("bottomright", paste("Cluster", levels(as.factor(final_tight[-whRm]))), fill=cc, cex=1)

pdf("DAT/tsne.pdf")
plot(tsne_data[idx,], t='p', main = "tSNE visualization (1,000 most variable genes)", col=cc[as.factor(final_tight[-whRm][idx])], pch = 19, xlab="Component 1", ylab="Component 2")
legend("topright", paste("Cluster", levels(as.factor(final_tight[-whRm]))), fill=cc, cex=.5)
dev.off()
```

```{r tsne_500, dependson="final"}
whRm <- which(final_tight == -1)
vars <- apply(correctedY_dat[,-whRm], 1, var)
vars <- sort(vars, decreasing=TRUE)

d <- dist(t(correctedY_dat[,-whRm][names(vars)[1:500],]))

set.seed(123)
tsne_data <- tsne(d, k = 2, perplexity = 5, max_iter = 2000)
```

```{r tsne_plot_500, dependson=c("tsne_500", "final"), results="markup"}
idx <- apply(tsne_data, 1, function(x) max(abs(x))<=200)

pdf("DAT/tsne_500.pdf")
plot(tsne_data[idx,], t='p', main = "tSNE visualization (500 most variable genes)", col=cc[as.factor(final_tight[-whRm][idx])], pch = 19, xlab="Component 1", ylab="Component 2")
legend("topright", paste("Cluster", levels(as.factor(final_tight[-whRm]))), fill=cc, cex=.5)
dev.off()
```

```{r tsne_250, dependson="final"}
whRm <- which(final_tight == -1)
vars <- apply(correctedY_dat[,-whRm], 1, var)
vars <- sort(vars, decreasing=TRUE)

d <- dist(t(correctedY_dat[,-whRm][names(vars)[1:250],]))

set.seed(123)
tsne_data <- tsne(d, k = 2, perplexity = 5, max_iter = 2000)
```

```{r tsne_plot_250, dependson=c("tsne_250", "final"), results="markup"}
idx <- apply(tsne_data, 1, function(x) max(abs(x))<=200)

pdf("DAT/tsne_250.pdf")
plot(tsne_data[idx,], t='p', main = "tSNE visualization (250 most variable genes)", col=cc[as.factor(final_tight[-whRm][idx])], pch = 19, xlab="Component 1", ylab="Component 2")
legend("topright", paste("Cluster", levels(as.factor(final_tight[-whRm]))), fill=cc, cex=.5)
dev.off()
```

# Comparison to Poulin

```{r poulin, dependson="final"}
poulin1 <- c("Gapdh", "Th", "Slc6a3", "Slc18a2", "Ddc", "Prkcd", "Grin2c", "Sncg", "Ndnf", "Igf1", "Sox6", "Aldh1a1", "Calb1", "Slc17a6", "Cck", "Foxa2", "Lmx1a", "Egr1", "Tacr3", "Grp", "Lpl", "Otx2", "Adcyap1", "Nefm", "Vip", "Chrna6", "Chrna4", "Syt17", "Ntn1", "Slc32a1", "Clstn2", "Zfhx3", "Gsg1l", "Satb1", "Clic5", "Cartpt", "Bmp2", "Kcns3", "Fgf1", "Gad1", "Fzd1", "Lrrk2", "Clic3", "B4galt1", "Aldh2", "Abcc8", "Nrip3", "Park2", "Shh", "Znf746", "Slc6a1", "Kcnj11", "Etv5", "Gucy2c", "Parm1", "Fam184a", "Gpr161", "Pitx3", "Cacna1d", "Lmx1b", "Cacna1c", "Calb2", "En1", "Syt1", "Slc10a4", "En2", "Nr4a2", "Snca", "Sdc2", "Dlk1", "Ntf3", "Vgf", "Cpne2", "Slit2", "Cntnap4", "Pink1", "Grik3", "Drd2", "Foxa1", "Gch1", "Ntsr1", "Grb10", "Efnb3", "Rab3c", "Kcnj6", "Lrrc3b", "Sema6d", "Zdhhc2", "Oxr1", "Bag1", "Marcks", "Clptm1l", "Park7", "Atp13a2")
poulin1 <- intersect(poulin1, rownames(fq_dat))

pdf("DAT/heatmap_poulin1.pdf")
dualHeatmap(clusterVector=final_tight,heatData=t(fq_dat),clusterData=dendro,clusterSamples=TRUE,whVars=poulin1,annCol=data.frame(FinalCluster=as.factor(final_tight),PAM=clusters_dat, Batch=batch_dat),main="Manually Curated Genes", annColors=list(FinalCluster=c("white", cc)))
dev.off()

poulin2 <- c("Sox6", "Sncg", "Ndnf", "Igf1", "Foxa2", "Lmx1a", "Aldh1a1", "Slc32a1", "Satb1", "Clstn2", "Adcyap1", "Lpl", "Otx2", "Vip", "Chrna4", "Gsg1l", "Snca", "Ntf3")

poulin2 <- intersect(poulin2, rownames(fq_dat))

pdf("DAT/heatmap_poulin2.pdf")
dualHeatmap(clusterVector=final_tight,heatData=t(fq_dat),clusterData=dendro,clusterSamples=TRUE,whVars=poulin2,annCol=data.frame(FinalCluster=as.factor(final_tight),PAM=clusters_dat, Batch=batch_dat),main="Manually Curated Genes", annColors=list(FinalCluster=c("white", cc)))
dev.off()

```

```{r manual_gene_hm, dependson="genesDE", results="markup", eval=FALSE}
genes1 <- c("Eif2ak2", "Flna", "Kctd12b", "Id2", "Magt1", "Cflar")
genes2 <- c("Tac1", "Nxph1", "Arpp21", "Irs4", "Adcyap1", "Rasgrp1", "Ar")
genes3 <- c("Egr1", "Gas7", "Fstl1", "Cd63", "Sparc", "Lpar1")
genes4 <- c("Tox", "Spock3", "Pkib", "Calcr", "Rab3b", "Arhgap36")
genes5 <- c("Ret", "Ntsr1", "Bsn", "Thy1", "Dynll2.1",
            "Syt1", "Ywhag", "Pink1", "Rnf208", "Megf9", "Gpr153", "Cplx1")
genes6 <- c("Sparcl1", "Gad2", "Crhbp", "Vcan", "Plcl1")
genes7 <- c("Ldhb", "Atp2b2", "Anxa1", "Aldh1a7", "Igf1", "Lrrtm4")
genes8 <- c("Neurod6", "Gkn1", "Grp", "Gpr83", "Lpl", "Atp1b2", "Gabra1", "Tcf4")
genes9 <- c("Ddah1", "Slc7a3", "Pcp4l1",
            "Fgf10", "Fam114a1", "Cbln1")

geneord <- c(genes6, genes4, genes5, genes2, genes9, genes7, genes8, genes1, genes3)

groupLength <- c(length(genes6), length(genes4), length(genes5), length(genes2), length(genes9), length(genes7), length(genes8), length(genes1), length(genes3))
sep <- cumsum(groupLength)

blockColors <- outDendoHeat$annColors$FinalCluster[which(!outDendoHeat$annColors$FinalCluster %in% c("white"))]

dendro<-clusterHclust(pcaDat_dat[-whRm,],final_tight[-whRm],full=TRUE,unassigned="remove")

out<-dualHeatmap(clusterVector=final_tight[-whRm],heatData=t(fq_dat)[geneord,-whRm],clusterData=dendro,clusterSamples=TRUE,clusterVars=TRUE,annCol=data.frame(FinalCluster=final_tight[-whRm]),whVars=1:nrow(fq_dat[geneord,-whRm]),
annColors=list(FinalCluster=blockColors,Cluster=blockColors),main="Interesting Genes",breaks=0.95,cexRow=1)

#function to add blank lines of NA in data of arbitrary size
makeBlankData<-function(data,sep,nadd=1){
  naData<-matrix(NA,nrow=nadd,ncol=ncol(data))
  colnames(naData)<-colnames(data)
	start<-c(1,sep+1)
	end<-c(sep,nrow(data))
	len<-end-start+1
	grFac<-rep(1:length(len),times=len)
	dataList<-by(data,grFac,function(x){x})
	dataListMinus<-lapply(dataList[-length(dataList)],function(x){
		return(rbind(x,naData))
	})
	rnames<-lapply(dataList,rownames)
	rnamesMinus<-lapply(head(rnames,-1),function(x){c(x,rep("",nadd))})
	rnames<-c(unlist(rnamesMinus),rnames[[length(rnames)]])
	return(list(data=do.call("rbind",c(dataListMinus,dataList[length(dataList)])),rownames=rnames))
}

names(final_tight) <- colnames(fq_dat)
cll <- factor(final_tight[-whRm], levels=c(6, 4, 5, 2, 9, 7, 8, 10, 1, 3))
cl <- sort(cll)
keep <- names(cl)
temp <- makeBlankData(fq_dat[geneord,keep],sep=head(sep,-1),nadd=2)
dat <- temp$data
rnames <- temp$rownames

cols <- outDendoHeat$annColors$FinalCluster
colors <- cols[-1][c(6, 4, 5, 2, 9, 7, 8, 10, 1, 3)]

pdf("DAT/manual_genes_heatmap.pdf",width=6,height=6)
dualHeatmap(clusterVector=cl,heatData=dat,clusterData=dat,clusterSamples=FALSE,clusterVars=FALSE,annCol=data.frame(Cluster=cl),whVars=1:nrow(dat),
annColors=list(colors),main="",breaks=out$breaks,cexRow=1,
treeheight=50,labCol=rep("",ncol(dat)),labRow=rnames)
dev.off()

```
